---
title: "p8105_revisions_hw3_dmb2257"
author: "Diane Benites"
date: "2025-11-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
This loads necessary packages and the dataset. 
```{r}
library(p8105.datasets)
library(tidyverse)
data("instacart")
```
Description of the Instacart Dataset: There are 1384617 observations of 15 variables. Each column represents one of the variables in the dataset. Each row represents one product of an order. This data provides information on when instacart orders in the sample were placed, which products were ordered, the frequency of orders, and the aisles associated with the orders. 


```{r}
items_ordered_df =
instacart|>
  janitor::clean_names()|>
  group_by(aisle_id, aisle)|>
  summarize(
    item_order_count=n())|>
  arrange(desc(item_order_count))
  
```
There are 134 aisles. The most items are ordered from aisle 83: fresh vegetables (150609 items ordered), aisle 24: fresh fruits (150473 items ordered), and aisle 123: packaged vegatables (78493 items ordered).

This makes a plot of the number of items ordered. The aisles are sorted by the number of items ordered. The names of the aisles are included on the x axis. 
```{r}
items_ordered_df|>
  filter(item_order_count >= 10000)|>
  ggplot(
    aes(fct_reorder(aisle, item_order_count), item_order_count))+
  geom_col()+
  labs(
    x = "Aisle Name",
    y = "Number of Items Ordered"
  )+
  theme(axis.text.x = element_text(angle = 70, hjust =1))
  
```

This creates a table of the 3 most popular items in the aisles "baking ingredients", "dog food care", "packaged vegetables fruits", and how many of each item was ordered.
```{r}
instacart|>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits"))|>
  group_by(aisle, product_name)|>
  summarize(
    product_count = n()
  )|>
  arrange(desc(product_count))|>
  filter(min_rank(desc(product_count))< 4)|>
  knitr::kable()
```

This makes a table showing the mean hour of the day at which pank lady apples and coffee ice cream were ordered.
```{r}
instacart |>
  mutate(
    order_dow = 
      case_match(
        order_dow,
        0 ~ "Sunday",
        1 ~ "Monday",
        2 ~ "Tuesday",
        3 ~ "Wednesday",
        4 ~ "Thursday",
        5 ~ "Friday",
        6 ~ "Saturday"),
    order_dow = as.factor(order_dow))|>
  group_by(product_name, order_dow)|>
  summarize(
    mean_hour = mean(order_hour_of_day, na.rm = TRUE)
  )|>
  filter(
    product_name %in% c("Pink Lady Apples", "Coffee Ice Cream"))|>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour)|>
  knitr::kable()
```

# Problem 2
This imports the datasets. It cleans the second dataset to create a month variable instead of the detailed date. It also removed missing price data points. 
```{r}
zipcodes_tidy_df = 
    read_csv("p8105_revisions_hw3_dmb2257_files/Zip Codes.csv", na = c("NA", ".", ""))|>
  janitor::clean_names()

zori_df = 
  read_csv("p8105_revisions_hw3_dmb2257_files/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv")|>
  pivot_longer(
    -(RegionID:CountyName),
    names_to = "dates",
    values_to = "price"
  )|>
  janitor::clean_names()|>
  rename(zip_code = region_name) |>
  mutate(
    dates = as_date(dates),
    zip_code = as.numeric(zip_code)
  )|>
  drop_na(price)

```


This code chunk summarizes the number of months each zip code was observed. 
```{r}
zip_count=
zori_df|>
    mutate(
    month = floor_date(dates, unit = "month")
  )|>
  select(-dates)|>
  group_by(zip_code)|>
  summarize(
    zip_obs = n_distinct(month)
  )

zip_count|>
  filter(zip_obs == 116)

zip_count |>
  filter(zip_obs < 10)


```
There were 48 zip codes observed 116 times. There were 26 zip codes observed fewer than 10 times.

```{r}
average_price_df =
zori_df|>
  mutate(
    year = floor_date(dates, unit = "year"))|>
  rename(
    borough = county_name)|>
  separate(year, into=c("year", "month", "day"))|>
  select(-month, -day, -dates)


average_price_df|>
  group_by(borough, year)|>
  summarize(
    avg_price = mean(price)
  )|>
  pivot_wider(
    names_from = borough,
    values_from = avg_price
  )|>
  knitr::kable()
```
Among all boroughs, the average rental price increases each year, with the exception of an decrease in average price in 2020 in Kings, NY County and Queens and in 2021 in Kings and Queens. These differences may be attributed to changes in rent the covid-19 pandemic.

```{r}
average_price_df|>
  group_by(zip_code, borough, year)|>
  summarize(price)|>
  ggplot(aes(x= year, y = price, color = borough)) +
  geom_point()

```


```{r}
```



# Problem 3


