---
title: "p8105_revisions_hw2_dmb2257"
author: "Diane Benites"
date: "2025-11-30"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Necessary Packages and Sets Data Path
```{r}
library(tidyverse)
library(janitor)
library(readxl)

data_path <- 
"C:/Users/dmben/OneDrive/Desktop/data_science_1/p8105_revisions_dmb2257"
```


## Problem 1

This code chunk loads the unemployment dataframe  and cleans the names. It selects the needed columns and mutates the months to the full name. 
```{r}
unemployment_df = 
  read_csv(file = "./unemployment.csv")|>
  janitor::clean_names() |>
  
  select(year, jan, feb, mar, apr, may, jun, jul, aug, sep, oct, nov, dec) |>
  
  pivot_longer(
    jan:dec, 
    names_to = "month",
    values_to = "percentages") |>
    
  mutate(
    month = 
      case_match(
       month,
       "jan" ~ "January",
       "feb" ~ "February",
       "mar" ~ "March",
       "apr" ~ "April",
       "may" ~ "May",
       "jun" ~ "June",
       "jul" ~ "July",
       "aug" ~ "August",
       "sep" ~ "September",
       "oct" ~ "October",
       "nov" ~ "November",
       "dec" ~ "December"),
     month = as.factor(month))

```

This code chunk imports the polsmonth data. It separates the date into year, month and date and renames the numeric value of the month to the full month name. It also selects all variables except for day, prez_gop and prez_dem. It also filters the data to only include the points that were yes fo dem or gop.
```{r}
polsmonth_df = 
  read_csv(file = "./pols-month.csv")|>

  janitor::clean_names() |>
  
  separate(mon, into = c("year", "month", "day")) |>

  mutate(
    month = 
      case_match(
       month,
       "01" ~ "January",
       "02" ~ "February",
       "03" ~ "March",
       "04" ~ "April",
       "05" ~ "May",
       "06" ~ "June",
       "07" ~ "July",
       "08" ~ "August",
       "09" ~ "September",
       "10" ~ "October",
       "11" ~ "November",
       "12" ~ "December"),
     month = as.factor(month)) |>
    select(year, month, prez_gop, prez_dem, gov_gop, gov_dem, sen_gop, sen_dem, rep_gop, rep_dem) |>
    
  pivot_longer(
    prez_gop:prez_dem,
    names_to = "president") |>
    
  mutate(
    president = case_match(
      president,
      "prez_gop" ~ "gop",
      "prez_dem" ~ "dem")) |>
    
  filter(value == 1)
```


The loads the dataset snp and separates the sate into month, day and year. It mutates the numeric month to the full month name. 
```{r}
snp_df = 
  read_csv(file = "./snp.csv")|>
janitor::clean_names() |>

separate(date, into = c("month", "day", "year")) |>

  mutate(
    month = 
      case_match(
       month,
       "01" ~ "January",
       "02" ~ "February",
       "03" ~ "March",
       "04" ~ "April",
       "05" ~ "May",
       "06" ~ "June",
       "07" ~ "July",
       "08" ~ "August",
       "09" ~ "September",
       "10" ~ "October",
       "11" ~ "November",
       "12" ~ "December"),
     month = as.factor(month)) |>
    relocate(year, month, close)|>
    select(-day)
```

The polsmonth dataset contains observations regarding the political parties of politicians. The "presidents" variable identifies the president during that given time as democrat or republican, demonstrated by value 1. The unemployment dataset contains observations regarding the percentage of unemployment during a given month each year. There were originally 13 variables, one for year and one for each of the 12 months. The months variables and values were consolidated into one month column and one percentages column, respectively. The snp dataset contains observations with 2 variables, the date of the observation and the closing S&P value at the given date. The date separated into month, date, and year, and the numeric values for month were renamed to the full month name. 


## Problem 2
This imports the mr. trashwheel data sheet and cleans the names to snakecase.
```{r}
mr_trashwheel_df <-
  read_excel("./202409 Trash Wheel Collection Data.xlsx", range = "Mr. Trash Wheel!A2:N655")|>
  janitor::clean_names()|>
  drop_na(dumpster)|>
  mutate(sports_balls = as.integer(sports_balls))

```


This imports and cleans  the names to snakecase in the Prof Trash Wheel data sheet.
```{r}
prof_trashwheel_df <-
  read_excel("./202409 Trash Wheel Collection Data.xlsx", range = "Professor Trash Wheel!A2:M123")|>
  janitor::clean_names()|>
  drop_na(dumpster)

```


This imports and cleans the names to snakecase in the Gwynnda data sheet.
```{r}
gwynnda_trashwheel_df <-
  read_excel("./202409 Trash Wheel Collection Data.xlsx", range = "Gwynnda Trash Wheel!A2:L266")|>
  janitor::clean_names()|>
  drop_na(dumpster)
```

This combines the datasets
```{r}
tidy_trashwheel_df = 
  left_join(mr_trashwheel_df, prof_trashwheel_df, by = "dumpster")

tidy_trashwheel_df2 =
  left_join(tidy_trashwheel_df, gwynnda_trashwheel_df, by = "dumpster")
```


## Problem 3
This imports and cleans the names to snakecase in the zip codes and zori data sets
```{r}
zipcodes_tidy_df = 
    read_csv(file = "./Zip Codes.csv")|>
  janitor::clean_names()

zori_df = 
  read_csv(file =  "./Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv")|>
  janitor::clean_names()


```

